---
df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

## Urban Climate Risk Assessment (UCRA)

Climate-risk indicators from the UCRA pipeline. UCRA runs as a **country batch**,
so the charts below place this city against its peer cities in the same run
(highlighted in red). Figures live under `03-render-output/ucra/`; the underlying
tables carry the `_ucra` suffix (cross-city tables are prefixed `ucra_stats_`).

Maps follow the cartography of the CRP *Urban Climate Risk Analysis — Burundi*
report: washed-out basemap, raster clipped to the AOI, black dashed AOI outline,
scale bar in kilometres.

```{r}
#| label: setup-ucra
#| include: false
if (file.exists("_quarto.yml")) here::i_am("tasks/ucra/charts/index.qmd")
if (!exists("aoi")) {
  source(here::here("core/R/setup.R"))
}
```

```{r}
#| label: data-loading-ucra
#| include: false

ucra_dir <- here::here("03-render-output", "ucra")
ucra_ls  <- function(sub) {
  d <- file.path(ucra_dir, sub)
  if (dir.exists(d)) list.files(d, pattern = "\\.png$", full.names = TRUE) else character(0)
}
ucra_maps  <- ucra_ls("maps")
ucra_plots <- ucra_ls("plots")
ucra_cmip6 <- ucra_ls("cmip6")

# Figures this task renders itself, plus the ones UCRA's own R scripts draw
# (r_scripts/*.R -> <project_dir>/plots/, exported here with an `r_` prefix).
# Anything not listed still appears, captioned by its file stem — so new output
# shows up without editing this file.
ucra_caption_map <- c(
  "lst_summer_ucra"              = "Land surface temperature — summer",
  "lst_winter_ucra"              = "Land surface temperature — winter",
  "wsf_evolution_ucra"           = "Built-up area evolution (WSF)",
  "air_quality_2019_ucra"        = "Air quality (PM2.5), 2019",
  "builtup_ssp2_2050_ucra"       = "Built-up projection — SSP2, 2050",
  "builtup_ssp3_2080_ucra"       = "Built-up projection — SSP3, 2080",
  "gdp_ucra"                     = "Gridded GDP",
  "urbanheat_ssp2_day_sum_ucra"  = "Urban heat island — SSP2, daytime",
  "urbanheat_ssp2_nig_sum_ucra"  = "Urban heat island — SSP2, night-time",
  "urbanheat_ssp5_day_sum_ucra"  = "Urban heat island — SSP5, daytime",
  "drought_twsan_ucra"           = "Terrestrial water storage anomaly",
  "avg_temp_peers_ucra"          = "Mean land surface temperature vs peer cities",
  "air_quality_trend_ucra"       = "Air quality (PM2.5), 1998-2019, vs peer-city mean",
  "cckp_anomaly_peers_ucra"      = "CCKP climate anomaly vs peer cities",
  "r_aep_ucra"                   = "Annual exceedance probability",
  "r_avg_temp_ucra"              = "Mean temperature, all cities",
  "r_avg_air_1998_2019_ucra"     = "Mean PM2.5, 1998-2019, all cities",
  "r_bad_air_1998_2019_ucra"     = "Days above the PM2.5 guideline, 1998-2019",
  "r_cdd_clim_ucra"              = "Consecutive dry days — climatology",
  "r_cdd_anom_ucra"              = "Consecutive dry days — anomaly",
  "r_hd35_tr26_clim_ucra"        = "Hot days (>35°C) and tropical nights (>26°C) — climatology",
  "r_hd35_tr26_anom_ucra"        = "Hot days and tropical nights — anomaly",
  "r_landslide_ucra"             = "Landslide hazard",
  "r_pr_clim_ucra"               = "Precipitation — climatology",
  "r_pr_anom_ucra"               = "Precipitation — anomaly",
  "r_r20mm_r50mm_clim_ucra"      = "Heavy rainfall days (>20mm, >50mm) — climatology",
  "r_r20mm_r50mm_anom_ucra"      = "Heavy rainfall days — anomaly",
  "r_r95ptot_clim_ucra"          = "Extreme rainfall contribution (R95pTOT) — climatology",
  "r_r95ptot_anom_ucra"          = "Extreme rainfall contribution — anomaly",
  "r_spei01_ucra"                = "SPEI-1 (short-term drought)",
  "r_spei12_ucra"                = "SPEI-12 (long-term drought)",
  "r_spei12_clim_ucra"           = "SPEI-12 — climatology",
  "r_spei48_ucra"                = "SPEI-48 (multi-year drought)",
  "r_tas_clim_ucra"              = "Air temperature — climatology",
  "r_tas_anom_ucra"              = "Air temperature — anomaly",
  "r_tas_txx_clim_ucra"          = "Air temperature and annual maximum (TXx) — climatology",
  "r_tas_txx_anom_ucra"          = "Air temperature and TXx — anomaly",
  "r_twsan_ucra"                 = "Terrestrial water storage anomaly",
  "r_wsdi_clim_ucra"             = "Warm spell duration index — climatology",
  "r_wsdi_anom_ucra"             = "Warm spell duration index — anomaly"
)

ucra_gallery <- function(files, heading) {
  if (length(files) == 0) return(invisible(NULL))
  cat("\n\n### ", heading, "\n\n", sep = "")
  stems   <- tools::file_path_sans_ext(basename(files))
  ordered <- c(intersect(names(ucra_caption_map), stems),
               sort(setdiff(stems, names(ucra_caption_map))))
  for (stem in ordered) {
    f <- files[stems == stem][1]
    if (is.na(f)) next
    cap <- if (!is.null(ucra_caption_map[[stem]])) ucra_caption_map[[stem]] else stem
    cat("\n\n#### ", cap, "\n\n", sep = "")
    cat(sprintf("![%s](%s)\n\n", cap, normalizePath(f, winslash = "/", mustWork = FALSE)))
  }
}

ucra_read <- function(pattern) {
  f <- list.files(tabular_dir, pattern = pattern, full.names = TRUE)[1]
  if (!is.na(f) && file.exists(f)) {
    tryCatch(read_csv(f, col_types = cols()), error = function(e) NULL)
  } else NULL
}
ucra_builtup  <- ucra_read("built_up_stats_ucra\\.csv$")
ucra_avg_temp <- ucra_read("^ucra_stats_avg_temp_ucra\\.csv$")
ucra_cmip6_tb <- ucra_read("cmip6_summary_ucra_cmip6\\.csv$")
```

```{r}
#| label: ucra-figures
#| results: asis

if (length(c(ucra_maps, ucra_plots, ucra_cmip6)) == 0) {
  print_text(paste_bold("No UCRA figures available — run `scan ucra` for this city."))
} else {
  ucra_gallery(ucra_maps,  "Maps")
  ucra_gallery(ucra_plots, "Peer-city comparisons and climate indices")
  ucra_gallery(ucra_cmip6, "CMIP6 downscaled projections (this city)")
}
```

```{r}
#| label: ucra-tables
#| results: asis

if (!is.null(ucra_builtup)) {
  cat("\n\n### Built-up area statistics (this city)\n\n")
  print(htmltools::tagList(ucra_builtup %>% print_paged_df()))
  cat("\n\n")
}
if (!is.null(ucra_cmip6_tb)) {
  cat("\n\n### CMIP6 projected change by scenario (2025 to 2100)\n\n")
  print(htmltools::tagList(ucra_cmip6_tb %>% print_paged_df()))
  cat("\n\n")
}
if (!is.null(ucra_avg_temp)) {
  cat("\n\n### Mean land surface temperature, all cities in the batch\n\n")
  print(htmltools::tagList(ucra_avg_temp %>% print_paged_df()))
  cat("\n\n")
}
if (is.null(ucra_builtup) && is.null(ucra_avg_temp) && is.null(ucra_cmip6_tb)) {
  print_text(paste_bold("No UCRA tables available."))
}
```
