---
df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

## Sectoral GDP Flood Exposure

High GDP areas are defined as pixels above the 60th percentile of GDP values within the AOI.

```{r}
#| label: setup-gdp-sectoral
#| message: false
#| warning: false
#| results: hide
if (file.exists("_quarto.yml")) here::i_am("tasks/gdp_sectoral/charts/index.qmd")
if (!exists("aoi")) {
  source(here::here("core/R/setup.R"))
}
```

```{r}
#| label: data-loading-gdp-sectoral
#| message: false
#| warning: false
#| results: hide
flood_gdp_sect <- tryCatch(
  read_csv(file.path(tabular_dir, "flood_gdp_sectoral.csv"), show_col_types = FALSE),
  error = function(e) tibble(flood_type = character(), sector = character(),
                             year = integer(), pct_high_gdp_in_flood = numeric()))

sector_colors <- c(Agriculture = "#2ca02c", Industry = "#ff7f0e", Service = "#800000")
sector_linetypes <- c(Agriculture = "solid", Industry = "dashed", Service = "dotted")

.plot_flood_sector <- function(ft_data, ft) {
  p <- ggplot(ft_data, aes(x = year, y = pct_high_gdp_in_flood, color = sector, linetype = sector, group = sector)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous(breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, NA), expand = expansion(c(0, 0.1)),
                       labels = scales::label_percent(scale = 1)) +
    scale_color_manual(values = sector_colors) +
    scale_linetype_manual(values = sector_linetypes) +
    theme_minimal() +
    labs(title = paste("High GDP areas per sector exposed by", ft, "flooding"),
         x = "", y = "% of high GDP pixels in flood zone", color = NULL, linetype = NULL) +
    theme(
      axis.line = element_line(linewidth = .5, color = "black"),
      legend.position = "bottom")
  ggsave(here("03-render-output", "plots", paste0("sectoral-gdp-flood-", tolower(ft), ".png")),
         plot = p, device = "png", width = 6, height = 4, units = "in", dpi = "print")
  p
}
```

### River Flooding

```{r}
#| label: chart-gdp-river-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "River", sector != "Combined")
if (nrow(ft_data) > 0) {
  print(.plot_flood_sector(ft_data, "River"))
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, sector, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No river flood data available"))
}
```

### Rainwater Flooding

```{r}
#| label: chart-gdp-rainwater-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "Rainwater", sector != "Combined")
if (nrow(ft_data) > 0) {
  print(.plot_flood_sector(ft_data, "Rainwater"))
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, sector, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No rainwater flood data available"))
}
```

### Coastal Flooding

```{r}
#| label: chart-gdp-coastal-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "Coastal", sector != "Combined")
if (nrow(ft_data) > 0) {
  print(.plot_flood_sector(ft_data, "Coastal"))
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, sector, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No coastal flood data available"))
}
```

## Total Sectoral GDP Flood Exposure

Total sectoral GDP is the sum of agriculture, industry, and service GDP per pixel. High GDP areas below are defined using the 60th percentile of this total.

### River Flooding

```{r}
#| label: chart-gdp-combined-river-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "River", sector == "Combined")
if (nrow(ft_data) > 0) {
  p <- ggplot(ft_data, aes(x = year, y = pct_high_gdp_in_flood, group = 1)) +
    geom_line(linewidth = 0.8, color = "#543005") +
    geom_point(size = 2, color = "#543005") +
    scale_x_continuous(breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, NA), expand = expansion(c(0, 0.1)),
                       labels = scales::label_percent(scale = 1)) +
    theme_minimal() +
    labs(title = "High total GDP areas exposed by River flooding",
         x = "", y = "% of high GDP pixels in flood zone") +
    theme(axis.line = element_line(linewidth = .5, color = "black"))
  ggsave(here("03-render-output", "plots", "combined-gdp-flood-river.png"), plot = p,
         device = "png", width = 6, height = 4, units = "in", dpi = "print")
  print(p)
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No river flood data available"))
}
```

### Rainwater Flooding

```{r}
#| label: chart-gdp-combined-rainwater-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "Rainwater", sector == "Combined")
if (nrow(ft_data) > 0) {
  p <- ggplot(ft_data, aes(x = year, y = pct_high_gdp_in_flood, group = 1)) +
    geom_line(linewidth = 0.8, color = "#543005") +
    geom_point(size = 2, color = "#543005") +
    scale_x_continuous(breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, NA), expand = expansion(c(0, 0.1)),
                       labels = scales::label_percent(scale = 1)) +
    theme_minimal() +
    labs(title = "High total GDP areas exposed by Rainwater flooding",
         x = "", y = "% of high GDP pixels in flood zone") +
    theme(axis.line = element_line(linewidth = .5, color = "black"))
  ggsave(here("03-render-output", "plots", "combined-gdp-flood-rainwater.png"), plot = p,
         device = "png", width = 6, height = 4, units = "in", dpi = "print")
  print(p)
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No rainwater flood data available"))
}
```

### Coastal Flooding

```{r}
#| label: chart-gdp-combined-coastal-flood
#| message: false
#| warning: false
#| results: asis
ft_data <- flood_gdp_sect %>% filter(flood_type == "Coastal", sector == "Combined")
if (nrow(ft_data) > 0) {
  p <- ggplot(ft_data, aes(x = year, y = pct_high_gdp_in_flood, group = 1)) +
    geom_line(linewidth = 0.8, color = "#543005") +
    geom_point(size = 2, color = "#543005") +
    scale_x_continuous(breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, NA), expand = expansion(c(0, 0.1)),
                       labels = scales::label_percent(scale = 1)) +
    theme_minimal() +
    labs(title = "High total GDP areas exposed by Coastal flooding",
         x = "", y = "% of high GDP pixels in flood zone") +
    theme(axis.line = element_line(linewidth = .5, color = "black"))
  ggsave(here("03-render-output", "plots", "combined-gdp-flood-coastal.png"), plot = p,
         device = "png", width = 6, height = 4, units = "in", dpi = "print")
  print(p)
  cat("\n\n")
  print_paged_df(ft_data %>% select(year, pct_high_gdp_in_flood, exposed_gdp_M, pct_gdp_exposed))
} else {
  print_text(paste_bold("No coastal flood data available"))
}
```

